---
title: "Homework 1"
author: Pallavi Krishnamurthy
output: html_document
---

This is my solution to Homework one.

```{r lib_calls}
library(tidyverse)
```

# problem 1 

## Part 1 :

Create a data frame with specified elements 
-a random sample of size 10 from a standard Normal distribution
-a logical vector indicating whether elements of the sample are greater than 0
-a character vector of length 10
-a factor vector of length 10, with 3 different factor “levels”

```{r dataframe}
prob1_df =
tibble(
  samp = rnorm(10),
  samp_gt_0 = samp > 0, 
  char_vec = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  factor_vec = factor(c("low","low", "low", "mod", "mod", "mod", "high", "high", "high", "high"))
)
```

Take the mean of each variable in the data frame prob1_df.

```{r dataset_mean}
mean(pull(prob1_df, samp)) 
mean(pull(prob1_df, samp_gt_0))
mean(pull(prob1_df, char_vec))
mean(pull(prob1_df, factor_vec))
```

Answer - The mean of a random sample of size 10 from a standard Normal distribution and the logical vector greater than zero gives a numeric result that is correct. However, the mean of the character vector and factor vector does not give a numeric results because it does not make sense to take the mean of "a", "b", "c", ... or "low", "mod" and "high". Hence, R gives a warning message that the argument if not numeric or logical : returning NA (Not Applicable). 

Explicitly convert variables from one type to another. Here we convert variables to numeric type. 

```{r covert_type_1, warning=FALSE, results='hide'}
as.numeric(pull(prob1_df, samp)) 
as.numeric(pull(prob1_df, samp_gt_0)) 
as.numeric(pull(prob1_df, char_vec)) 
as.numeric(pull(prob1_df, factor_vec)) 
```

Answer - The results for character vector is still not applicable. This is correct. The results for factor variable now yields a numeric value. This is because R forces the factor vector to take a numeric value when we explicitly change the variable type to numeric. We have to be cautious while changing types to make sure the results make sense. 

## Part2:

convert the logical vector to numeric, and multiply the random sample by the result
convert the logical vector to a factor, and multiply the random sample by the result
convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result

```{r covert_type_2}
as.numeric(pull(prob1_df, samp_gt_0)) * pull(prob1_df, samp)
as.factor(pull(prob1_df, samp_gt_0)) * pull(prob1_df, samp)
as.numeric(as.factor(pull(prob1_df, samp_gt_0))) * pull(prob1_df, samp)
```

Answer - We can see that factor vector simply multiplied by a numeric value does not yield a numeric result. (This is correct). However, when we change the type to numeric we get a numeric result that may not be correct. 

# Problem 2 

```{r dataset_penguins}
data("penguins", package = "palmerpenguins")
```

The penguins dataset has the the following variable names - `r names(penguins)`. The species are `r levels(penguins$species)` with `r count(filter(penguins, species == 'Adelie'))`, `r count(filter(penguins, species == 'Chinstrap'))` and `r count(filter(penguins, species == 'Gentoo'))` number of penguins for each species, respectively, and the islands are namely `r levels(penguins$island)` with `r count(filter(penguins, island == 'Biscoe'))`, `r count(filter(penguins, island == 'Dream'))` and `r count(filter(penguins, island == 'Torgersen'))` number of penguins for each island, respectively. The average bill length is `r mean(penguins$bill_length_mm, na.rm = TRUE)` and bill depth is `r mean(penguins$bill_depth_mm, na.rm = TRUE)`. The average body mass is `r mean(penguins$body_mass_g, na.rm = TRUE)`. The dataset has `r ncol(penguins)` variables and `r nrow(penguins)` observations. The mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)`. 

## Scatterplot flipper length vs bill length. 

```{r scatterplot}
penguin_length_plot = ggplot(data = penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +
  xlab("Bill Length (mm)") + ylab("Flipper Length (mm)")
  penguin_length_plot

ggsave("penguin_length_plot.pdf")
```

